Fixing Vulnerabilities Hinders Maintainability

### 2020-01-28 paper/icpc20.tex
patterns table

  of our datasets are collected before the analysis performed by BCH.s

@@

yielded a database of $624$ patches~\cite{10.1109/MSR.2019.00064}. The original dataset contains $1282$ commits. One patch can have multiple commits assigned.

@@

used to patch the vulnerability. We assumed that the last commit of the patch as the fix (\emph{sha}) and the parent of the first commit as the vulnerable version (\emph{sha-p}).

@@

Enumeration IDs (CWE). The most prevailing categories are the following (categories
with less than $20$ instances were merged in a bigger category named \textit{Miscellaneous}):
    \item \textbf{CWE-89: Improper Neutralization of Special Elements used in an SQL Command/SQL Injection (23 cases).} When developers do not keep untrusted data
    separate from SQL queries. If an attacker sends a command that
    exploits the syntax of the SQL interpreter, then SQL injection attack is possible.
    \item \textbf{CWE-611: Improper Restriction of XML External Entity Reference (24 cases).} 
	XML documents are exchanged through the web containing entities with URIs
	that resolve to local/external files. Thus, when XML parsers are not well configured, 
	attackers have allowed to directly those files.
    \item \textbf{CWE-20: Improper Input Validation (76 cases).} When developers
	forget to validate the inputs of an application, attackers may have control
	of the control/data flow of the program.
    \item \textbf{CWE-200: Information Exposure (39 cases).} Application's logs are one way of intentional/unintentional disclosure information
	to attackers. Many times attackers get access to logs when sould not be authorized to.
	\item \textbf{CWE-352: Cross-Site Request Forgery/CSRF (22 cases).} Poor session tokens generation
    and management usually allow attackers to send forged HTTP requests
    including authentication information from the victim to the vulnerable web
    application.
	\item \textbf{CWE-264: Permissions, Privileges, and Access Controls (59 cases).} Incorrectly
	    implemented functionalities related to authentication and session
	    management, allowing an attacker to gain access to session tokens,
	    passwords, keys, and other sensitive data.
	\item \textbf{CWE-399: Resource Management Errors (22 cases).} Resource management issues are found more frequently
		in programming languages that do not manage memory automatically (e.g., C/C++
	    and Objective-C), i.e., where developers are responsible for
	    handling it instead. It is one of the main causes of DoS attacks.
	\item \textbf{CWE-79: Improper Neutralization of Input During Web Page Generation/Cross-site Scripting (62 cases)} Lack of proper validation or escaping
	    allows attackers to submit untrusted data to web browsers through malicious
	    scripts that can hijack the user sessions or redirect the user to malicious
	    sites.
	\item \textbf{CWE-22: Improper Limitation of a Pathname to a Restricted Directory/Path Traversal (37 cases).}  When developers forget to neutralize special elements within the pathnames to files/directories, attackers 
	may leverage this flaw to resolve to a location outside of the restricted directory and gain access to documents
	or entire repositories.
	\item \textbf{Miscellaneous (673 cases)} This pattern comprises
		several other security refactorings that do not have a CWE assigned and patterns that
		do not satisfy the Wilcoxon test size requirement of more than $20$ refactorings.
\subsection{Security vs. Baseline Commits}
Previous studies attempted to measure the impact of regular refactorings on
open-source software maintainability~\cite{HEGEDUS2018313}. However, there is no
previous work focused on comparing the impact of security refactorings with
regular refactorings on maintainability.
We analyze the maintainability of regular commits (i.e., commits not related
to security fixes) and use them as a baseline.

@@

\begin{table}[H]
	\caption{Guidelines to produce maintainable code}
\begin{tabular}{L{1.66cm}L{3.25cm}L{2.65cm}}

>>>>

  of our datasets are collected before the analysis performed by BCH.

@@

yielded a database of $624$ patches~\cite{10.1109/MSR.2019.00064}. The original 
dataset contains $1282$ commits. One patch can have multiple commits assigned.

@@

used to patch the vulnerability. We assumed that the last commit of the patch as 
the fix (\emph{sha}) and the parent of the first commit as the vulnerable version 
(\emph{sha-p}).

@@

Enumeration IDs (CWE). In Table~\ref{tab:patterns}, the most prevailing categories are presented. Patterns
with less than $20$ instances were merged in a bigger pattern named \textit{Miscellaneous}.

\begin{table}[h]
\footnotesize
	\caption{Guidelines to produce maintainable code}
\begin{tabular}{L{1.1cm}L{2.1cm}L{4.3cm}}

\toprule
\textbf{Pattern} & \textbf{Name} & \textbf{Description}\\
\midrule

\textbf{CWE-89} & 
 Improper Neutralization of Special Elements used in an SQL Command/SQL Injection (23 cases) & 
 When developers do not keep untrusted data
     separate from SQL queries. If an attacker sends a command that
     exploits the syntax of the SQL interpreter, then SQL injection attack is possible.\\\midrule

\textbf{CWE-611} & 
 Improper Restriction of XML External Entity Reference (24 cases) & 
	XML documents are exchanged through the web containing entities with URIs
 	that resolve to local/external files. Thus, when XML parsers are not well configured,
 	attackers have allowed to directly those files.\\\midrule

 \textbf{CWE-20} & 
 Improper Input Validation (76 cases) & 
	When developers
 	forget to validate the inputs of an application, attackers may have control
 	of the control/data flow of the program.\\\midrule

  \textbf{CWE-200} & 
Information Exposure (39 cases)& 
Application's logs are one way of intentional/unintentional disclosure information
 	to attackers. Many times attackers get access to logs when sould not be authorized to.\\\midrule

   \textbf{CWE-352} & 
Cross-Site Request Forgery/CSRF (22 cases)& 
Poor session tokens generation
     and management usually allow attackers to send forged HTTP requests
     including authentication information from the victim to the vulnerable web
     application.\\\midrule

       \textbf{CWE-264} & 
Permissions, Privileges, and Access Controls (59 cases)& 
Incorrectly
 	    implemented functionalities related to authentication and session
 	    management, allowing an attacker to gain access to session tokens,
 	    passwords, keys, and other sensitive data.\\\midrule

 	           \textbf{CWE-399} & 
Resource Management Errors (22 cases)& 
Resource management issues are found more frequently
 		in programming languages that do not manage memory automatically (e.g., C/C++
 	    and Objective-C), i.e., where developers are responsible for
 	    handling it instead. It is one of the main causes of Denial-of-Service attacks.\\\midrule

 	     	           \textbf{CWE-79} & 
Improper Neutralization of Input During Web Page Generation/Cross-site Scripting (62 cases)& 
Lack of proper validation or escaping
 	    allows attackers to submit untrusted data to web browsers through malicious
 	    scripts that can hijack the user sessions or redirect the user to malicious
 	    sites.\\\midrule

 	     	           \textbf{CWE-22} & 
Improper Limitation of a Pathname to a Restricted Directory/Path Traversal (37 cases) & 
When developers forget to neutralize special elements within the path names to files/directories, attackers
 	may leverage this flaw to resolve to a location outside of the restricted directory and gain access to documents
 	or entire repositories.\\\midrule

 	 	     	           \textbf{MISC} & 
Miscellaneous (673 cases) & 
This pattern comprises
 		several other security refactorings that do not have a CWE assigned and patterns that
 		do not satisfy the Wilcoxon test size requirement of more than $20$ refactorings.\\
\bottomrule
\end{tabular}
\label{tab:patterns}
\end{table}

%
% \begin{itemize}
%     \item \textbf{CWE-89: Improper Neutralization of Special Elements used in an SQL Command/SQL Injection (23 cases).} When developers do not keep untrusted data
%     separate from SQL queries. If an attacker sends a command that
%     exploits the syntax of the SQL interpreter, then SQL injection attack is possible.
% %
%     \item \textbf{CWE-611: Improper Restriction of XML External Entity Reference (24 cases).}
% 	XML documents are exchanged through the web containing entities with URIs
% 	that resolve to local/external files. Thus, when XML parsers are not well configured,
% 	attackers have allowed to directly those files.
% %
%     \item \textbf{CWE-20: Improper Input Validation (76 cases).} When developers
% 	forget to validate the inputs of an application, attackers may have control
% 	of the control/data flow of the program.
% %
%     \item \textbf{CWE-200: Information Exposure (39 cases).} Application's logs are one way of intentional/unintentional disclosure information
% 	to attackers. Many times attackers get access to logs when sould not be authorized to.
% %
% 	\item \textbf{CWE-352: Cross-Site Request Forgery/CSRF (22 cases).} Poor session tokens generation
%     and management usually allow attackers to send forged HTTP requests
%     including authentication information from the victim to the vulnerable web
%     application.
% %
% 	\item \textbf{CWE-264: Permissions, Privileges, and Access Controls (59 cases).} Incorrectly
% 	    implemented functionalities related to authentication and session
% 	    management, allowing an attacker to gain access to session tokens,
% 	    passwords, keys, and other sensitive data.
% %
% 	\item \textbf{CWE-399: Resource Management Errors (22 cases).} Resource management issues are found more frequently
% 		in programming languages that do not manage memory automatically (e.g., C/C++
% 	    and Objective-C), i.e., where developers are responsible for
% 	    handling it instead. It is one of the main causes of DoS attacks.
% %
% 	\item \textbf{CWE-79: Improper Neutralization of Input During Web Page Generation/Cross-site Scripting (62 cases)} Lack of proper validation or escaping
% 	    allows attackers to submit untrusted data to web browsers through malicious
% 	    scripts that can hijack the user sessions or redirect the user to malicious
% 	    sites.
% %
% 	\item \textbf{CWE-22: Improper Limitation of a Pathname to a Restricted Directory/Path Traversal (37 cases).}  When developers forget to neutralize special elements within the pathnames to files/directories, attackers
% 	may leverage this flaw to resolve to a location outside of the restricted directory and gain access to documents
% 	or entire repositories.
% %
% 	\item \textbf{Miscellaneous (673 cases)} This pattern comprises
% 		several other security refactorings that do not have a CWE assigned and patterns that
% 		do not satisfy the Wilcoxon test size requirement of more than $20$ refactorings.
% \end{itemize}
%
\subsection{Security vs. Baseline Commits}
%
Previous studies attempted to measure the impact of regular refactorings on
open-source software maintainability~\cite{HEGEDUS2018313}. However, there is no
previous work focused on comparing the impact of security refactorings with
regular refactorings on maintainability.
We analyze the maintainability of regular commits (i.e., commits not related
to security fixes) and use them as a baseline.

@@

\begin{table}[h]
\footnotesize
	\caption{Guidelines to produce maintainable code}
\begin{tabular}{L{1.66cm}L{3.3cm}L{2.7cm}}
